---
title: "Group Project - Did covid effect Asian hate crime? "
author: "Wan Ting Hsu"
date: "2/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

## Prepare data for model
# Load packages
```{r}
library(dplyr)
library(tidyverse)
library(lubridate)
library(readr)
library(hrbrthemes)
library(bsts)
library(ggplot2)
library(patchwork)
```

#Read data 
```{r}

nypd <- read.csv("nypd.csv")

```

```{r}
#change date format
nypd <- nypd %>% mutate(CMPLNT_FR_DT = mdy(CMPLNT_FR_DT ))
nypd$CMPLNT_FR_DT<-as.Date(nypd$CMPLNT_FR_DT)  

#filter the date period we want (from 2010 March)
new<- nypd %>% 
  filter(CMPLNT_FR_DT >= "2010-03-01") %>%
  dplyr::select(CMPLNT_FR_DT, VIC_RACE) %>%
  group_by(CMPLNT_FR_DT, VIC_RACE) %>%
  count(VIC_RACE)

#group the data to monthly
new$CMPLNT_FR_DT <- format(as.Date(new$CMPLNT_FR_DT), "%Y-%m")
colnames(new)[1] = "date"
colnames(new)[2] = "race"
str_trim(new$race, side = "both")

#delete missing data
new <- new[-which(new$race == ""), ]

```


#Create before & after covid
```{r}
#bc = 10 years period before covid 
bc <- new %>%
  filter(date <"2020-03") %>%
  group_by(date) %>%
  count(race) 

#ac = after covid from 2020-03 to 2021-12-31
ac <- new %>%
  filter(date >= "2020-03") %>%
  group_by(date) %>%
  count(race)
```



#pivot_wider
```{r}
ac <- ac %>% pivot_wider(names_from = race, values_from = n)
bc <- bc %>% pivot_wider(names_from = race, values_from = n)

ac[is.na(ac)] = 0
bc[is.na(bc)] = 0
```

#create asian/non-asian
```{r}
#group all the non-asian race to one variable of "non_asian"
ac$non_asian <- rowSums(ac[,c("BLACK", "BLACK HISPANIC", "UNKNOWN", "WHITE", "WHITE HISPANIC", "AMERICAN INDIAN/ALASKAN NATIVE")])
bc$non_asian <- rowSums(bc[,c("BLACK", "BLACK HISPANIC", "UNKNOWN", "WHITE", "WHITE HISPANIC", "AMERICAN INDIAN/ALASKAN NATIVE")])
```

```{r}
ac <- ac %>% dplyr::select(date, `ASIAN / PACIFIC ISLANDER`, non_asian)
bc <- bc %>% dplyr::select(date, `ASIAN / PACIFIC ISLANDER`, non_asian)
```



#before covid correlation
We want to find out if there is a correlation between Asian hate crime and Non-asian hate crime using before covid data. The reason for this is because if there is a high correlation between Asian and Non-asian hate crime, we can predict the after covid hate crime ratio using before covid hate crime data. 
```{r}

cor(bc$`ASIAN / PACIFIC ISLANDER`, bc$non_asian)

```

The result shows a 0.79 high correlation between asiancrime and non-asian crime. We are confidence to derive the indication that other race  crime is highly correlated with asian crime. With this in mind, we can want to use bsts time series model to predict after covid asian crime using other racecrime and compare to actual asian crime data. 




## Time Series

#Asian Crime to all crime Ratio
We created a asian crime ratio (asian crime/all crime) in order to see weather if the percentage of Asian crime is difference with our time series prediction. 
```{r}
ac$ratio = round(ac$`ASIAN / PACIFIC ISLANDER`/sum(ac$`ASIAN / PACIFIC ISLANDER`,ac$non_asian),5)
bc$ratio = round(bc$`ASIAN / PACIFIC ISLANDER`/sum(bc$`ASIAN / PACIFIC ISLANDER`,bc$non_asian),5)
```




#change dataframe to xts format

```{r}

library(xts)

ac$date<-as.yearmon(ac$date)
ac <- xts(ac$ratio, ac$date)
bc$date<-as.yearmon(bc$date)
bc <- xts(bc$ratio, bc$date)
  
colnames(ac)[1] = "ratio"
colnames(bc)[1] = "ratio"

```


#bsts model (bc ratio) 
```{r}
library(bsts)     # load the bsts package
library(zoo)

ss<- AddLocalLinearTrend(list(), bc$ratio)
ss <- AddSeasonal(ss, bc$ratio, nseasons = 12)
model1 <- bsts(bc$ratio,
               state.specification = ss,
               niter = 1000,
               seed = 5205)
```


#plot bc ratio

Plot before covid 
```{r}
plot(model1)
plot(model1, "components")
plot(model1, "help")


```
#Predict using bc ratio

```{r}
#jpeg('BTST_Prediction.jpg')

pred1 <- predict(model1, horizon =  22, seed = 5205)
plot(pred1, 
     font = 4,
     main = "BSTS Prediction",
     par(bg = "#f7f7f7"),
     xlab = "Time", ylab = "Crime Ratio", 
     ylim = c(0, 0.015), 
     plot.original = 60)

#dev.off()

```

```{r}
pred1$interval
```


```{r}
plot(pred1, 
     font = 4,
     main = "BSTS Prediction",
     par(bg = "#f7f7f7"),
     xlab = "Time", ylab = "Crime Ratio", 
     ylim = c(0, 0.015), 
     plot.original = FALSE)
```


# ploting the actual ratio after-covid
```{r}
#jpeg('actual.jpg')

plot(ac,     
     font = 4,
     col="red",
     par= "white",
     main = "Actual After Covid Crime Ratio",
     par(bg = "#f7f7f7"),
     xlab = "Time", ylab = "Crime Ratio", 
     ylim = c(0, 0.015)
     )

#dev.off()
```



#Combine plot bc predtion & ac data
Here we combine the crime ratio of the after covid asian crime ratio prediction using before covid data and the actual after covid data to see if there is a difference. If we see a major difference, we can assume that covid might have some effects on asian  crime. However, if the two graphs are really similar, then it indicates that covid had no effect on asian  crime.

#get prediction data ready
```{r}
pred_df <- data.frame(date = c("2020-03","2020-04", "2020-05", "2020-06","2020-07", "2020-08","2020-09", "2020-10","2020-11", "2020-12" , "2021-01", "2021-02", "2021-03", "2021-04", "2021-05", "2021-06", "2021-07", "2021-08", "2021-09", "2021-10", "2021-11", "2021-12"), ratio = pred1$mean)
pred_df$ratio = round(pred_df$ratio,5)

pred_df$date<-as.yearmon(pred_df$date)
pred_df <- xts(pred_df$ratio, pred_df$date)
colnames(pred_df)[1] = "ratio"
```

#combine plot ac & bc prediction

```{r}
#jpeg('combine.jpg')

plot.zoo(cbind(ac, pred_df), 
         plot.type = "single", 
         col = c("red", "blue"),
         main = "Combination",
          xlab = "Time", ylab = "Asian Crime Ratio")
legend("bottomright", 
       legend = c("Actual Crime Rate","Prediction"),
       col = c("red","blue"),
       lwd = 1)
#dev.off()
```



##Quantify Result

#t-distribution
H0: E(Y1) - E(Y1^) = 0
HA: E(Y1) - E(Y1^) ≠ 0
E(Y1): After covid Asian  crime ratio
E(Y1^): Pre covid prediction for after covid Asian crime ratio
With 95% confidence interval


According to the t-test result, our p-value > 0.05 hence we accept our H0 hypothesis, which is there is no difference in the mean towards Asian crime pre-covid and after-covid. Covid-19 did not increase Asian crime within New York City. 
```{r}
t.test(pred_df$ratio,ac$ratio)
```







