---
title: "Group Project - Did covid effect Asian hate crime? "
author: "Wan Ting Hsu"
date: "2/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

## Prepare data for model
# Load packages
```{r}
library(dplyr)
library(tidyverse)
library(lubridate)
library(readr)
library(hrbrthemes)
library(bsts)
library(ggplot2)
library(patchwork)
library(ggthemes)
library(gridExtra) 
library(quantmod);
library(xts);
library(zoo) 
library(forecast) 
library(fpp); 
library(fpp2) 
library(tseries) 

```

#Read data 
```{r}

nypd <- read.csv("nypd.csv")

```

```{r}
#change date format
nypd <- nypd %>% mutate(CMPLNT_FR_DT = mdy(CMPLNT_FR_DT ))
nypd$CMPLNT_FR_DT<-as.Date(nypd$CMPLNT_FR_DT)  

#filter the date period we want (from 2010 March)
new<- nypd %>% 
  filter(CMPLNT_FR_DT >= "2010-03-01") %>%
  dplyr::select(CMPLNT_FR_DT, VIC_RACE) %>%
  group_by(CMPLNT_FR_DT, VIC_RACE) %>%
  count(VIC_RACE)

#group the data to monthly
new$CMPLNT_FR_DT <- format(as.Date(new$CMPLNT_FR_DT), "%Y-%m")
colnames(new)[1] = "date"
colnames(new)[2] = "race"
str_trim(new$race, side = "both")

#delete missing data
new <- new[-which(new$race == ""), ]

```


#Create before & after covid
```{r}
#bc = 10 years period before covid 
bc <- new %>%
  filter(date <"2020-03") %>%
  group_by(date) %>%
  count(race) 

#ac = after covid from 2020-03 to 2021-12-31
ac <- new %>%
  filter(date >= "2020-03") %>%
  group_by(date) %>%
  count(race)
```



#pivot_wider
```{r}
ac <- ac %>% pivot_wider(names_from = race, values_from = n)
bc <- bc %>% pivot_wider(names_from = race, values_from = n)

ac[is.na(ac)] = 0
bc[is.na(bc)] = 0
```

#create asian/non-asian
```{r}
#group all the non-asian race to one variable of "non_asian"
ac$all<- rowSums(ac[,c("ASIAN / PACIFIC ISLANDER","BLACK", "BLACK HISPANIC", "UNKNOWN", "WHITE", "WHITE HISPANIC", "AMERICAN INDIAN/ALASKAN NATIVE")])
bc$all <- rowSums(bc[,c("ASIAN / PACIFIC ISLANDER","BLACK", "BLACK HISPANIC", "UNKNOWN", "WHITE", "WHITE HISPANIC", "AMERICAN INDIAN/ALASKAN NATIVE")])
```

#lm1
We want to find out if there is a correlation between Asian hate crime and Non-asian hate crime using before covid data. The reason for this is because if there is a high correlation between Asian and Non-asian hate crime, we can predict the after covid hate crime ratio using before covid hate crime data. 

```{r}
lm1 = lm(`ASIAN / PACIFIC ISLANDER` ~ BLACK + `BLACK HISPANIC` + UNKNOWN + WHITE +`WHITE HISPANIC` + `AMERICAN INDIAN/ALASKAN NATIVE` + all, data = bc)
summary(lm1)
```

#all crime
After looking at the correlation of the every other race to asian crime, we find high correlation for all of them. Hence, we decided to use all crime and asian crime as the base for our analysis. 

We created a crime ratio (race crime/all crime) in order to see weather if the percentage of Asian crime is difference with our time series prediction. 

```{r}
#Asian
ac_a <- ac %>% dplyr::select(date, `ASIAN / PACIFIC ISLANDER`, all)
bc_a <- bc %>% dplyr::select(date, `ASIAN / PACIFIC ISLANDER`, all)

ac_a$ratio = round((ac_a$`ASIAN / PACIFIC ISLANDER`/ac_a$all),3)
bc_a$ratio = round((bc_a$`ASIAN / PACIFIC ISLANDER`/bc_a$all),3)
```

```{r}
#Black
ac_b <- ac %>% dplyr::select(date, BLACK, all)
bc_b <- bc %>% dplyr::select(date, BLACK, all)

ac_b$ratio = round((ac_b$BLACK/ac_b$all),3)
bc_b$ratio = round((bc_b$BLACK/bc_b$all),3)
```

```{r}
#Hispanic
ac_bh <- ac %>% dplyr::select(date, `BLACK HISPANIC`, `WHITE HISPANIC`, all)
bc_bh <- bc %>% dplyr::select(date, `BLACK HISPANIC`, `WHITE HISPANIC` , all)

ac_h <- ac_bh %>% mutate(HISPANIC = sum(`BLACK HISPANIC`,`WHITE HISPANIC`)) 
bc_h <- bc_bh %>% mutate(HISPANIC = sum(`BLACK HISPANIC`,`WHITE HISPANIC`)) 

ac_h$ratio = round((ac_h$HISPANIC/ac_h$all),3)
bc_h$ratio = round((bc_h$HISPANIC/bc_h$all),3)


```

```{r}
#White
ac_w <- ac %>% dplyr::select(date, WHITE, all)
bc_w <- bc %>% dplyr::select(date, WHITE, all)

ac_w$ratio = round((ac_w$WHITE/ac_w$all),3)
bc_w$ratio = round((bc_w$WHITE/bc_w$all),3)
```



```{r}
#AMERICAN INDIAN/ALASKAN NATIVE
ac_ai <- ac %>% dplyr::select(date, `AMERICAN INDIAN/ALASKAN NATIVE`, all)
bc_ai <- bc %>% dplyr::select(date, `AMERICAN INDIAN/ALASKAN NATIVE`, all)

ac_ai$ratio = round((ac_ai$`AMERICAN INDIAN/ALASKAN NATIVE`/ac_ai$all),3)
bc_ai$ratio = round((bc_ai$`AMERICAN INDIAN/ALASKAN NATIVE`/bc_ai$all),3)
```



#before-covid correlation between Anti-Asian crime and all crime

```{r}

cor(bc$`ASIAN / PACIFIC ISLANDER`, bc$all)

```

The result of linear regression shows a high correlation between asian crime and all crime. With this in mind, we can use bsts time series model to predict after covid asian crime using other racecrime and compare to actual asian crime data. 




## Time Series

#change dataframe to xts format


```{r}

library(xts)

#Asian
ac_a$date<-as.yearmon(ac_a$date)
ac_a <- xts(ac_a$ratio, ac_a$date)
bc_a$date<-as.yearmon(bc_a$date)
bc_a <- xts(bc_a$ratio, bc_a$date)
  
colnames(ac_a)[1] = "ratio"
colnames(bc_a)[1] = "ratio"

#Black
ac_b$date<-as.yearmon(ac_b$date)
ac_b <- xts(ac_b$ratio, ac_b$date)
bc_b$date<-as.yearmon(bc_b$date)
bc_b <- xts(bc_b$ratio, bc_b$date)
  
colnames(ac_b)[1] = "ratio"
colnames(bc_b)[1] = "ratio"

#Hispanic
ac_h$date<-as.yearmon(ac_h$date)
ac_h <- xts(ac_h$ratio,ac_h$date)
bc_h$date<-as.yearmon(bc_h$date)
bc_h <- xts(bc_h$ratio, bc_h$date)
  
colnames(ac_h)[1] = "ratio"
colnames(bc_h)[1] = "ratio"

#White
ac_w$date<-as.yearmon(ac_w$date)
ac_w <- xts(ac_w$ratio, ac_w$date)
bc_w$date<-as.yearmon(bc_w$date)
bc_w <- xts(bc_w$ratio, bc_w$date)
  
colnames(ac_w)[1] = "ratio"
colnames(bc_w)[1] = "ratio"

#American Indian/Alaskan

ac_ai$date<-as.yearmon(ac_ai$date)
ac_ai <- xts(ac_ai$ratio, ac_ai$date)
bc_ai$date<-as.yearmon(bc_ai$date)
bc_ai <- xts(bc_ai$ratio, bc_ai$date)
  
colnames(ac_ai)[1] = "ratio"
colnames(bc_ai)[1] = "ratio"

```



#bsts model (bc ratio) 

```{r}
library(bsts)     # load the bsts package
library(zoo)

#Asian
ss_a<- AddLocalLinearTrend(list(), bc_a$ratio)
ss_a <- AddSeasonal(ss_a, bc_a$ratio, nseasons = 12)
model_a <- bsts(bc_a$ratio,
               state.specification = ss_a,
               niter = 1000,
               seed = 5205)

#Black
ss_b<- AddLocalLinearTrend(list(), bc_b$ratio)
ss_b <- AddSeasonal(ss_b, bc_b$ratio, nseasons = 12)
model_b <- bsts(bc_b$ratio,
               state.specification = ss_b,
               niter = 1000,
               seed = 5205)

#Hispanic
ss_h<- AddLocalLinearTrend(list(), bc_h$ratio)
ss_h <- AddSeasonal(ss_h, bc_h$ratio, nseasons = 12)
model_h <- bsts(bc_h$ratio,
               state.specification = ss_h,
               niter = 1000,
               seed = 5205)

#White
ss_w<- AddLocalLinearTrend(list(), bc_w$ratio)
ss_w <- AddSeasonal(ss_w, bc_w$ratio, nseasons = 12)
model_w <- bsts(bc_w$ratio,
               state.specification = ss_w,
               niter = 1000,
               seed = 5205)

#American-indian
ss_ai<- AddLocalLinearTrend(list(), bc_ai$ratio)
ss_ai <- AddSeasonal(ss_ai, bc_ai$ratio, nseasons = 12)
model_ai <- bsts(bc_ai$ratio,
               state.specification = ss_ai,
               niter = 1000,
               seed = 5205)


```


#plot bc ratio

Plot before covid 


```{r}
#Asian
#jpeg('actual1.jpg')

plot(model_a)

#dev.off()

```

```{r}
#jpeg('trend_seasonal.jpg')
plot(model_a, "components")
#dev.off()
```



```{r}
plot(model_a, "residuals")
```



```{r}
#Black
plot(model_b)
plot(model_b, "components")
plot(model_b, "help")


```
```{r}
#Hispanic
plot(model_h)
plot(model_h, "components")
plot(model_h, "help")


```

```{r}
#white
plot(model_w)
plot(model_w, "components")
plot(model_w, "help")


```

```{r}
#American Indian
plot(model_ai)
plot(model_ai, "components")
plot(model_ai, "help")


```

#Predict using bc ratio



```{r}
#Asian

#jpeg('BTST_Prediction1.jpg')

pred_a <- predict(model_a, horizon =  22, seed = 5205)
plot(pred_a, 
     font = 4,
     main = "BSTS Prediction",
     par(bg = "#f7f7f7"),
     xlab = "Time", ylab = "Crime Ratio", 
     ylim = c(-0.2, 0.4), 
     plot.original = 60)

#dev.off()

```

```{r}
#Black

#jpeg('BTST_Prediction1.jpg')

pred_b <- predict(model_b, horizon =  22, seed = 5205)
plot(pred_b, 
     font = 4,
     main = "BSTS Prediction",
     par(bg = "#f7f7f7"),
     xlab = "Time", ylab = "Crime Ratio", 
     ylim = c(-0.4, 1), 
     plot.original = 60)

#dev.off()

```
```{r}
#Hispanic
#jpeg('BTST_Prediction1.jpg')

pred_h <- predict(model_h, horizon =  22, seed = 5205)
plot(pred_h, 
     font = 4,
     main = "BSTS Prediction",
     par(bg = "#f7f7f7"),
     xlab = "Time", ylab = "Crime Ratio", 
     ylim = c(-0.4, 1), 
     plot.original = 60)

#dev.off()

```

```{r}
#White
#jpeg('BTST_Prediction1.jpg')

pred_w <- predict(model_w, horizon =  22, seed = 5205)
plot(pred_w, 
     font = 4,
     main = "BSTS Prediction",
     par(bg = "#f7f7f7"),
     xlab = "Time", ylab = "Crime Ratio", 
     ylim = c(-0.3, 1), 
     plot.original = 60)

#dev.off()

```

```{r}
#American Indian
#jpeg('BTST_Prediction1.jpg')

pred_ai <- predict(model_ai, horizon =  22, seed = 5205)
plot(pred_ai, 
     font = 4,
     main = "BSTS Prediction",
     par(bg = "#f7f7f7"),
     xlab = "Time", ylab = "Crime Ratio", 
     ylim = c(-0.1, 1), 
     plot.original = 60)

#dev.off()

```




```{r}
#Look closly into BSTS prediction of Asian related crime 

#jpeg('BSTS_Pred_only.jpg')

plot(pred_a, 
     font = 4,
     main = "BSTS Prediction",
     par(bg = "#f7f7f7"),
     xlab = "Time", ylab = "Crime Ratio", 
     ylim = c(-0.2, 0.4), 
     plot.original = FALSE)


#dev.off()
```


# ploting the actual ratio after-covid


```{r}

#Asian 

#jpeg('actual1.jpg')

plot(ac_a,     
     font = 4,
     col="red",
     par= "white",
     main = "Actual After Covid Crime Ratio",
     par(bg = "#f7f7f7"),
     xlab = "Time", ylab = "Crime Ratio", 
     ylim = c(0, 0.2)
     )

#dev.off()
```

```{r}
#Black
#jpeg('actual1.jpg')

plot(ac_b,     
     font = 4,
     col="red",
     par= "white",
     main = "Actual After Covid Crime Ratio",
     par(bg = "#f7f7f7"),
     xlab = "Time", ylab = "Crime Ratio", 
     ylim = c(0, 0.4)
     )

#dev.off()
```

```{r}
#Hispanic
#jpeg('actual1.jpg')

plot(ac_h,     
     font = 4,
     col="red",
     par= "white",
     main = "Actual After Covid Crime Ratio",
     par(bg = "#f7f7f7"),
     xlab = "Time", ylab = "Crime Ratio", 
     ylim = c(0, 0.4)
     )

#dev.off()
```

```{r}
#White
#jpeg('actual1.jpg')

plot(ac_w,     
     font = 4,
     col="red",
     par= "white",
     main = "Actual After Covid Crime Ratio",
     par(bg = "#f7f7f7"),
     xlab = "Time", ylab = "Crime Ratio", 
     ylim = c(0, 0.4)
     )

#dev.off()
```

```{r}
#American Indian 
#jpeg('actual1.jpg')

plot(ac_ai,     
     font = 4,
     col="red",
     par= "white",
     main = "Actual After Covid Crime Ratio",
     par(bg = "#f7f7f7"),
     xlab = "Time", ylab = "Crime Ratio", 
     ylim = c(0, 0.4)
     )

#dev.off()
```

#Combine plot bc predtion & ac data
Here we combine the crime ratio of the after covid asian crime ratio prediction using before covid data and the actual after covid data to see if there is a difference. If we see a major difference, we can assume that covid might have some effects on asian  crime. However, if the two graphs are really similar, then it indicates that covid had no effect on asian  crime.

#get prediction data ready


```{r}
#Asian
pred_df_a <- data.frame(date = c("2020-03","2020-04", "2020-05", "2020-06","2020-07", "2020-08","2020-09", "2020-10","2020-11", "2020-12" , "2021-01", "2021-02", "2021-03", "2021-04", "2021-05", "2021-06", "2021-07", "2021-08", "2021-09", "2021-10", "2021-11", "2021-12"), ratio = pred_a$mean)
pred_df_a$ratio = round(pred_df_a$ratio,3)

pred_df_a$date<-as.yearmon(pred_df_a$date)
pred_df_a <- xts(pred_df_a$ratio, pred_df_a$date)
colnames(pred_df_a)[1] = "ratio"
```


```{r}
#Black
pred_df_b <- data.frame(date = c("2020-03","2020-04", "2020-05", "2020-06","2020-07", "2020-08","2020-09", "2020-10","2020-11", "2020-12" , "2021-01", "2021-02", "2021-03", "2021-04", "2021-05", "2021-06", "2021-07", "2021-08", "2021-09", "2021-10", "2021-11", "2021-12"), ratio = pred_b$mean)
pred_df_b$ratio = round(pred_df_b$ratio,3)

pred_df_b$date<-as.yearmon(pred_df_b$date)
pred_df_b <- xts(pred_df_b$ratio, pred_df_b$date)
colnames(pred_df_b)[1] = "ratio"
```

```{r}
#Hispanic
pred_df_h <- data.frame(date = c("2020-03","2020-04", "2020-05", "2020-06","2020-07", "2020-08","2020-09", "2020-10","2020-11", "2020-12" , "2021-01", "2021-02", "2021-03", "2021-04", "2021-05", "2021-06", "2021-07", "2021-08", "2021-09", "2021-10", "2021-11", "2021-12"), ratio = pred_h$mean)
pred_df_h$ratio = round(pred_df_h$ratio,3)

pred_df_h$date<-as.yearmon(pred_df_h$date)
pred_df_h <- xts(pred_df_h$ratio, pred_df_h$date)
colnames(pred_df_h)[1] = "ratio"
```

```{r}
#white
pred_df_w <- data.frame(date = c("2020-03","2020-04", "2020-05", "2020-06","2020-07", "2020-08","2020-09", "2020-10","2020-11", "2020-12" , "2021-01", "2021-02", "2021-03", "2021-04", "2021-05", "2021-06", "2021-07", "2021-08", "2021-09", "2021-10", "2021-11", "2021-12"), ratio = pred_w$mean)
pred_df_w$ratio = round(pred_df_w$ratio,3)

pred_df_w$date<-as.yearmon(pred_df_w$date)
pred_df_w <- xts(pred_df_w$ratio, pred_df_w$date)
colnames(pred_df_w)[1] = "ratio"
```

```{r}
#American Indian
pred_df_ai <- data.frame(date = c("2020-03","2020-04", "2020-05", "2020-06","2020-07", "2020-08","2020-09", "2020-10","2020-11", "2020-12" , "2021-01", "2021-02", "2021-03", "2021-04", "2021-05", "2021-06", "2021-07", "2021-08", "2021-09", "2021-10", "2021-11", "2021-12"), ratio = pred_ai$mean)
pred_df_ai$ratio = round(pred_df_ai$ratio,3)

pred_df_ai$date<-as.yearmon(pred_df_ai$date)
pred_df_ai <- xts(pred_df_ai$ratio, pred_df_ai$date)
colnames(pred_df_ai)[1] = "ratio"
```



#combine plot ac & bc prediction


```{r}
#Asian
#jpeg('combine1.jpg')

plot.zoo(cbind(ac_a, pred_df_a), 
         plot.type = "single", 
         col = c("red", "blue"),
         main = "Combination",
         ylim = c(0,0.2),
          xlab = "Time", ylab = "Asian Crime Ratio")
legend("bottomright", 
       legend = c("Actual Crime Rate","Prediction"),
       col = c("red","blue"),
       lwd = 1)
#dev.off()
```

```{r}
#Black
#jpeg('combine1.jpg')

plot.zoo(cbind(ac_b, pred_df_b), 
         plot.type = "single", 
         col = c("red", "blue"),
         main = "Combination",
         ylim = c(0,0.4),
          xlab = "Time", ylab = "Asian Crime Ratio")
legend("bottomright", 
       legend = c("Actual Crime Rate","Prediction"),
       col = c("red","blue"),
       lwd = 1)
#dev.off()
```


```{r}
#Hispanic
#jpeg('combine1.jpg')

plot.zoo(cbind(ac_h, pred_df_h), 
         plot.type = "single", 
         col = c("red", "blue"),
         main = "Combination",
         ylim = c(0,0.4),
          xlab = "Time", ylab = "Asian Crime Ratio")
legend("bottomright", 
       legend = c("Actual Crime Rate","Prediction"),
       col = c("red","blue"),
       lwd = 1)
#dev.off()
```

```{r}
#White
#jpeg('combine1.jpg')

plot.zoo(cbind(ac_w, pred_df_w), 
         plot.type = "single", 
         col = c("red", "blue"),
         main = "Combination",
         ylim = c(0,0.4),
          xlab = "Time", ylab = "Asian Crime Ratio")
legend("bottomright", 
       legend = c("Actual Crime Rate","Prediction"),
       col = c("red","blue"),
       lwd = 1)
#dev.off()
```

```{r}
#American-indian
#jpeg('combine1.jpg')

plot.zoo(cbind(ac_ai, pred_df_ai), 
         plot.type = "single", 
         col = c("red", "blue"),
         main = "Combination",
         ylim = c(0,0.4),
          xlab = "Time", ylab = "Asian Crime Ratio")
legend("bottomright", 
       legend = c("Actual Crime Rate","Prediction"),
       col = c("red","blue"),
       lwd = 1)
#dev.off()
```




##Quantify Result - Asian

We#t-distribution
H0: E(Y1) - E(Y1^) = 0
HA: E(Y1) - E(Y1^) ≠ 0
E(Y1): After covid Asian  crime ratio
E(Y1^): Pre covid prediction for after covid Asian crime ratio
With 95% confidence interval


According to the t-test result, our p-value < 0.05 hence we reject our H0 hypothesis, which is there is no difference in the mean towards Asian crime pre-covid and after-covid. Covid-19 indeed had an effect on Asian related crime in New York City. 
```{r}
t.test(pred_df_a$ratio,ac_a$ratio)
```






